### 2020-02-24 Common/chapters/introduction.tex
Julien and Johan: motivation

\section{Motivation}
xx

\section{Contributions}
This work proposes various methods to detect cancer on MRIs, with a focus on prostate cancer. Concretely, the major contributions of this work are:
\begin{itemize}

\item Processing scripts for the PROSTATEx dataset, the Lung CT Challenge dataset and the Kaggle Brain dataset. This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images multiple subsets (training, validation and test) that can be used as input to machine learning algorithms.

@@

\item Verification scripts to check the gradient flow of a neural network, the cropping of images, the presence of NaN value in images, etc. All these scripts can easily be imported in other projects.
\end{itemize}

>>>>

\section{Motivation}
Cancer is a serious concern in today's society. An early detection is key to maximize the chances of recovery. As explained by Nicholas Petrick~\cite{50}, "it has long been recognized that clinicians do not always make optimal use of the data acquired by an imaging device. The limitations of the human eye-brain system, limitations in training and experience, and factors such as fatigue, distraction, and satisfaction of search may all contribute to suboptimal use of available information". Under these circumstances, CAD systems can contribute to address some of these issues by being used as an aid for clinicians. In fact, they provide multiple advantages.

First of all, used as an extra diagnosis, they decrease the probability of missing positive diagnoses. Then, they can also speed up the diagnosis process by proposing regions of interest to clinicians, which, in time, could reduce the screening costs.\\
Similarly, lower screening costs imply a greater accessibility to screening tests, which can result in an earlier cancer detection. As stated before, detecting cancers at an early stage maximizes the probability of curing them.\\
Another benefit of using CAD systems, especially the ones based on deep learning, is the number of cases that the latter are trained on. In fact, in order to built efficient deep learning models, large amounts of data are required. Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.

However, these advantages remain partially theoretical nowadays. Research is still ongoing, aiming at improving existent CAD systems. Problems like poor detection rates, small amounts of available data and bad levels of generalization are still recurrent. Therefore, this situation can and must be improved.

%Consequently, this thesis presents how to build a modern Deep Learning model for prostate lesions classification and proposes a model based on transfer learning to overcome the lack of available data in the medical field.

\section{Contributions}
This work proposes a state of the art deep learning model for prostate lesion classification and proposes a method based on transfer learning to overcome the lack of publicly available data in the medical field. Concretely, the major contributions of this work are:
\begin{itemize}

\item Processing scripts for the PROSTATEx dataset, the Lung CT Challenge dataset and the Kaggle Brain dataset. This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images multiple subsets (training, validation and test) that can be used as input to machine learning algorithms. Also, the class imbalance problem is taked into account. The processing scripts automatically augment each class in order to balance the number of elements. In addition to this, an alternative relying on undersampling was implemented as a PyTorch sampler.

@@

\item Verification scripts to check the gradient flow of a neural network, the cropping of images, the presence of NaN values in images, etc. All these scripts can easily be used in other projects.
\end{itemize}
